### 2017-03-03 lafaySimulatedScenePerception.tex
archive old version

Results are shown on Table~\ref{tab:markers}. Nine markers are found across all semantic levels. As suggested by Figures~\ref{fig:xp1_class_1} and~\ref{fig:xp1_class_2}, sounds evoking human activity (\textit{male footsteps concrete}) and nature (\textit{animal}, \textit{birds} and \textit{birds singing}) are markers of the i-scene. The presence of \textit{church bell} and \textit{bell bike} as strong markers of the i-scenes can be due to the socio-cultural background of the subjects, in great majority European citizens. It confirms Schafer's claim that sounds that are recognized by a community as integral part of its sonic environment are popular \cite{schafer1977tuning}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{../gfxMatlab/pa5_1.eps}
\caption{\label{fig:pa5} Precision at rank 5 ($P@5$) computed from the \textit{Hamming} distances between the scenes at different semantic levels, considering the event and texture classes ($ET$), the event classes ($E$), the texture classes ($T$), the event markers ($E_m$) and event classes without event markers ($E_{w/o m}$) .}
\end{center}
\end{figure}

@@

Each simulated scene $s_i$ is represented by a boolean vector of dimension $n$: $S_i = (x_1, x_2, \ldots{}, x_{n})$, $i\in[1,72]$. Each dimension corresponds to a sound class (event and texture) of a particular semantic level (for example: $n=44$ classes for semantic level 1). Thus $x_1 = 1$ indicates that the sound class $x_1$ is present in scene $s_i$ (not present if $x_1=0$). A \textit{Hamming} distance is computed between all pairs of $S_i$. Then, for each vector $S_i$ the number of scenes sharing the same label (i or ni) as $S_i$  among the five scenes $s_j$ ($j=1,...,5$) that have the smallest distance to it is computed. The $P@5$ precision is then obtained by averaging the results over all the scenes $s_i$ ($i=1, \ldots{}, 72$).

Results considering both the event and texture classes ($ET$) \nm{symbolisme pas clair} \gl{pk ?} to build the vectors  $S_i$ are shown on Figure~\ref{fig:pa5}. The $P@5$ is of 81\% for semantic level 0 and remains above 89\% starting at semantic level 1. These result strongly indicates that it is possible to make a clear distinction between the i- and ni-scenes based solely on the class presence/absence. We also note that the deeper the hierarchical level, the higher the $P@5$; in other words an increased precision in identifying scene elements allows a better distinction between i- and ni-scenes.

To refine the analysis, only event classes ($E$), texture classes ($T$), event markers ($E_m$) and event classes without event markers ($E_{w/o m}$) are also considered on Figure~\ref{fig:pa5}. The constant deviation of circa $10\%$ from the semantic level 1 between $E$ and $T$, and the fact that $P@5$ are similar for $ET$ and $E$ indicate that soundscape quality is mostly conveyed by sound events.

Since similar precision are achieved by $ET$, $E$ and $E_m$, we may conclude that all the event classes are not equivalently important to distinguish between the two scene types. Indeed, $E_m$ uses very partial information but still reaches quasi-equivalent results. Moreover, the drop in $P@5$ considering $E_{w/o m}$ from the semantic level 1 confirms the fact that soundscape quality mostly relies on the presence of specific sound markers.

@@

In order to check if the physical characteristics of the markers alone have an influence on the pleasantness, we now consider the $D$ and the $L$ with on one hand the tracks of the markers ($L_m$, $D_{m}$) and on the other hand the remaining sounds ($L_b$, $D_b$). As the marker durations are rather short, and the number of markers in a scene are rather low, a window length of 125 ms is used to compute $L_m$ and $L_b$ \nm{pourquoi cette precision methodologique n'a t'elle pas ete donnee aussi au moment de definir ce parametre ?} \gl{plus simple ici}. For the i-scenes we consider only the markers of the i-scenes (i-markers), and for the ni-scenes only the markers of the ni-scenes (ni-markers). \nm{cela parait evident, pourquoi y'a-t'il besoin de le preciser ?} \gl{fixe bien les idées}

The difference $L_{m-b}=L_m-L_b$ is also considered as it illustrates how the markers emerge from the other sounds. Correlations between those descriptors and the average perceptual ratings are depicted in Table~\ref{tab:numMarkers} and Figure~\ref{fig:numMarkers}.

@@

Considering the ni-scenes, the good correlation for $L_m$  (Figure~\ref{fig:numMarkers}.b) and the absence of correlation for $L_{b}$  (Figure~\ref{fig:numMarkers}.c) suggests that the sound levels of the ni-markers convey most of the unpleasantness. Also, no correlation is observed for $L_{m-b}$ (Figure~\ref{fig:numMarkers}.a). This indicates that the predominant influence of the marker sound levels remains true irrespective of the sound level of the other sounds. A loud horn sound is expected to increase the annoyance level of a scene (more so than a soft klaxon sound), whether or not the remaining of the sonic environment is ``\,noisy\,''.

@@

These results are strong indicators that global sound level is not a good indicator to characterize the appraisal of \emph{a priori} positive sound environments such as a park or a quiet street. On the contrary, the contribution of each sound source needs to be considered separately, as for certain type of sounds, the pleasantness does increase with their sound levels.

>>>>

Results are shown on Table~\ref{tab:markers}. Nine markers are found across all semantic levels. As suggested by Figures~\ref{fig:xp1_class_1}.a, sounds evoking human activity (\textit{male footsteps concrete}) and nature (\textit{animal}, \textit{birds} and \textit{birds singing}) are markers of the i-scene. The presence of \textit{church bell} and \textit{bell bike} as strong markers of the i-scenes can be due to the socio-cultural background of the subjects, in great majority European citizens. It confirms Schafer's claim that sounds that are recognized by a community as integral part of its sonic environment are popular \cite{schafer1977tuning}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{../gfxMatlab/pa5_1.eps}
\caption{\label{fig:pa5} Precision at rank 5 ($P@5$) computed from the \textit{Hamming} distances between the scenes at different semantic levels, considering the event and texture classes ($ET$), the event classes ($E$), the texture classes ($T$), the event markers ($E_m$) and event classes without event markers ($E_{w/o,m}$) .}
\end{center}
\end{figure}

@@

Each simulated scene $s_i$ is represented by a boolean vector of dimension $n$: $S_i = (x_1, x_2, \ldots{}, x_{n})$, $i=\lbrace 1, \ldots, 72 \rbrace$. Each dimension corresponds to a sound class (event and texture) of a particular semantic level (for example: $n=44$ classes for semantic level 1). Thus $x_1 = 1$ indicates that the sound class $x_1$ is present in scene $s_i$ (not present if $x_1=0$). A \textit{Hamming} distance is computed between all pairs of $S_i$. Then, for each scene $s_i$, the number of scenes sharing the same label (i or ni) among the five scenes $s_j$ that have the smallest distance to $s_i$ is computed. The $P@5$ precision is then obtained by averaging the results over the 72 scenes $s_i$.

Results considering both the event and texture classes ($ET$) \nm{symbolisme pas clair} \gl{pk ?} to build the vectors  $S_i$ are shown on Figure~\ref{fig:pa5}. The $P@5$ is of 76\% for semantic level 0 and remains above 86\% starting at semantic level 1. These result strongly indicates that it is possible to make a clear distinction between the i- and ni-scenes based solely on the class presence/absence. We also note that the deeper the hierarchical level, the higher the $P@5$; in other words an increased precision in identifying scene elements allows a better distinction between i- and ni-scenes.

To refine the analysis, only event classes ($E$), texture classes ($T$), event markers ($E_m$) and event classes without event markers ($E_{w/o,m}$) are also considered on Figure~\ref{fig:pa5}. The constant deviation of circa $15\%$ from the semantic level 1 between $E$ and $T$, and the fact that $P@5$ are similar for $ET$ and $E$ indicate that soundscape quality is mostly conveyed by sound events.

Since similar precision are achieved by $ET$, $E$ and $E_m$, we may conclude that all the event classes are not equivalently important to distinguish between the two scene types. Indeed, $E_m$ uses very partial information but still reaches quasi-equivalent results. Moreover, the drop in $P@5$ considering $E_{w/o,m}$ from the semantic level 1 confirms the fact that soundscape quality mostly relies on the presence of specific sound markers.

@@

In order to check if the physical characteristics of the markers alone have an influence on the pleasantness, we now compute $D$ and $L$ considering on one hand the tracks of the markers ($L_m$, $D_{m}$) only, and on the other hand the remaining of the sounds ($L_b$, $D_b$). As the marker durations are rather short, and the number of markers in a scene rather low, a window length of 125 ms is used to compute $L_m$ and $L_b$ \nm{pourquoi cette precision methodologique n'a t'elle pas ete donnee aussi au moment de definir ce parametre ?} \gl{non, elle change ici}. The difference $L_{m-b}=L_m-L_b$ is also considered as it illustrates how the markers emerge from the other sounds. Correlations between those descriptors and the average ratings are depicted in Table~\ref{tab:numMarkers} and Figure~\ref{fig:numMarkers}.

@@

Considering the ni-scenes, the good correlation for $L_m$  (Figure~\ref{fig:numMarkers}.b) and the absence of correlation for $L_{b}$  (Figure~\ref{fig:numMarkers}.c) suggest that sound levels of the ni-markers convey most of the unpleasantness. The fact that no correlation is observed for $L_{m-b}$ (Figure~\ref{fig:numMarkers}.a) indicates that the predominant influence of the marker sound levels remains true irrespective of the sound level of the other sounds. A loud horn sound is expected to increase the annoyance level of a scene (more so than a soft klaxon sound), whether or not the remaining of the sonic environment is ``\,noisy\,''.

@@

These results are strong indicators that global sound level alone is not sufficient to properly characterize the appraisal of \emph{a priori} positive sound environments such as a park or a quiet street. On the contrary, the contribution of each sound source needs to be considered separately, as for certain type of sounds, the pleasantness does increase with their sound levels.
